Cost- and Deadline-Constrained Provisioning for Scientific Workflow Ensembles in IaaS Clouds

### 2013-07-12 paper/paper.tex
New plots for FGCS paper

\documentclass[preprint,12pt]{elsarticle}
%\documentclass[final,1p,times,twocolumn]{elsarticle}
\usepackage{fancyhdr}
\usepackage{graphicx}
\graphicspath{{./figures/}}
\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\usepackage[cmex10]{amsmath}
\usepackage{algorithm,algpseudocode}
\usepackage{hyperref} % must come before url
\usepackage{url}
\usepackage{color} % for \textcolor
\usepackage{subfig} % for \subfloat

@@

@@

Our work is related to the strategies for deadline-con\-strained
cost-minimization workflow scheduling, developed for utility grid systems. However, our problem is
different from \cite{Yu2005} and \cite{Abrishami2010} in that we consider ensembles of workflows in
IaaS clouds, which allow one to provision resources on a per-hour billing model,
rather than utility grids, which allow one to choose from a pool of existing
resources with a per-job billing model. Our work is also different from
cloud-targeted autoscaling solution~\cite{Mao2011} in that we consider ensembles
of workflows rather than unpredictable workloads containing workflows. We also consider budget constraints
rather than cost minimization as a goal. In other words, we assume that there is
more work to be done than the available budget, so some work must be rejected.
Therefore, cost is not something we optimize (i.e. an objective), but rather a constraint.

@@

\begin{figure}[tb] 
  \centering
  \subfloat[DPDS]{
  \includegraphics[width=0.40\textwidth]{figures/output-test-DPDS}
  \label{fig:algorithm-example-dpds}
  }\\
  \subfloat[SPSS]{
  \includegraphics[width=0.40\textwidth]{figures/output-test-SPSS}
  \label{fig:algorithm-example-spss}
  }
  \caption[Example schedules generated by the algorithms]{Example
  schedules generated by the algorithms. Each row is a different
  VM. Boxes are tasks colored by workflow.}

@@

\subsection{Relative Performance of Algorithms}

@@

\subsection{Task Granularity}

In the last experiment we noted that, for Montage and CyberShake, the short
runtimes of their tasks made the dynamic algorithms perform better relative to
SPSS. In order to test this theory we adjusted the granularity of the tasks in
several Montage and CyberShake ensembles to see how this would affect the
relative performance of the algorithms. The granularity adjustment was
achieved by multiplying the runtime of each task by a fixed scaling factor.

Fig.~\ref{fig:stretching} shows the relative performance of the algorithms
as the scaling factor is increased from 1 to 16. The figure only shows results
for the CyberShake application with Pareto unsorted ensembles, but similar
results were found for other ensemble types for both Montage and CyberShake.
Each data point in the figure represents 500 simulations (5 ensembles with 10
budgets and 10 deadlines). The best scores percent was calculated the same way
as it was for Fig.~\ref{fig:distributions}.

The figure shows that, as the scaling factor increases beyond 2, the relative
performance of SPSS surpasses that of the dynamic algorithms. Similar results
were found for other ensemble types for both applications. This result
suggests that, in general, for fine-grained workflows the dynamic algorithms
will produce better scores, and for coarse-grained workflows SPSS will produce
better scores.

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.3\textwidth]{stretching_cybershake}
    \caption{Percentage of high scores achieved by each algorithm for Pareto 
    unsorted CyberShake ensembles when task runtime is scaled by a constant 
    factor. A scaling factor of $x$ means that the runtime of tasks in each 
    workflow was multiplied by $x$.}
    \label{fig:stretching}
\end{figure}

>>>>

\documentclass[preprint,5p]{elsarticle}
%\documentclass[final,1p,times,twocolumn]{elsarticle}
\usepackage{fancyhdr}
\usepackage{graphicx}
\graphicspath{{./figures/}}
\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\usepackage[cmex10]{amsmath}
\usepackage{algorithm,algpseudocode}
\usepackage{hyperref} % must come before url
\usepackage{url}
\usepackage{color} % for \textcolor
\usepackage{subfig} % for \subfloat

@@

This paper is an extension of our earlier conference
publication~\cite{Malawski-SC12}.
Here we present a more detailed performance evaluation, including the discussion
of results based on deadline and budget constraints. We also report on new
results and discussion of task granularity in the context of the resource
billing cycle, which is important aspect of scheduling and resource provisioning
on cloud infrastructures.

@@

Our work is related to the strategies for deadline-con\-strained
cost-minimization workflow scheduling, developed for utility grid systems. However, our problem is
different from \cite{Yu2005} and \cite{Abrishami2013158} in that we consider ensembles of workflows in
IaaS clouds, which allow one to provision resources on a per-hour billing model,
rather than utility grids, which allow one to choose from a pool of existing
resources with a per-job billing model. Our work is also different from
cloud-targeted autoscaling solution~\cite{Mao2011} in that we consider ensembles
of workflows rather than unpredictable workloads containing workflows. We also consider budget constraints
rather than cost minimization as a goal. In other words, we assume that there is
more work to be done than the available budget, so some work must be rejected.
Therefore, cost is not something we optimize (i.e. an objective), but rather a constraint.

@@

\begin{figure}[tb] 
  \centering
  \subfloat[DPDS]{
  \includegraphics[width=0.80\columnwidth]{figures/output-test-DPDS}
  \label{fig:algorithm-example-dpds}
  }\\
  \subfloat[SPSS]{
  \includegraphics[width=0.80\columnwidth]{figures/output-test-SPSS}
  \label{fig:algorithm-example-spss}
  }
  \caption[Example schedules generated by the algorithms]{Example
  schedules generated by the algorithms. Each row is a different
  VM. Boxes are tasks colored by workflow.}

@@

\subsection{Relative Performance of Algorithms}
\label{sec:relative-performance}

@@

\subsection{Results vs. deadline and budget}

In order to give more details about the performance of the algorithms, in this
section we show how the results depend on the deadline and budget constraints.
As the performance metric we use the number of completed workflows of the
ensemble, since the exponential score is not suitable for showing trends.
We use all deadline and budget parameters as defined in
Section~\ref{sec:exp-parameters} and normalize them to the range from 0 to 1 to
facilitate plotting.
For each normalized deadline, we compute the mean number of completed workflows,
averaged over all budgets and random seeds, so each point on the plot is based
on 100 different simulation runs. In Fig.~\ref{fig:completed-deadlines}
and~\ref{fig:completed-budgets} we present the results of CyberShake and LIGO
applications, as representatives of fine-grained and coarse-grained workflows.
Similar results were obtained for other applications.

Fig.~\ref{fig:completed-deadlines} shows the results depending on normalized
deadline. It can be observed, that with increasing deadline, the number of
completed worklfows grows steeply and then stabilizes. This means that for
longer deadlines it is usually not possible to complete more workflows due to
the budget constraint. For short deadlines, the algorithms need to allocate more
VM instances in parallel, making it more difficult to utilize the resources
efficiently, which results in lower numbers of completed workflows.

We can also observe that the algorithms are able to complete more workflows in
the ensembles that have priorities assigned randomly (unsorted), in comparison
to the ensembes where the priorities are assigned based on workflow size
(sorted). This results from the fact that in the unsoted distributions the large
workflows often have low priorities, so they can be rejected by the algoritms,
thus making more space in the schedule for smaller high-priority workflows.

One interesting observation from Fig.~\ref{fig:completed-deadlines} is that for
CyberShake application the result curves have a sawtooth shape. This results
from the combination of two effects. The first comes from the fact that dynamic
algorithms perform better when the deadline is close to the full hour, due to
the provisioning mechanism that tries to shut down unused VM instances before a
full hour. The second one is that when the deadline increases, it may become
possible to admit a larger workflow with higher priority instead of several
smaller ones with lower priority. That is why the number of completed workflows
may decrese, but the exponential score (utility function) based on priorities
will be higher. These effects are visible for CyberShake application, for which
the maximum deadline is 3 hours, but they become less important for LIGO
application, for which the maximum deadline is 20 hours. See also the discussion
on task granularity in Setion~\ref{sec:granularity}.

Fig~\ref{fig:completed-budgets} shows the results based on normalized budget,
each point being the average over all deadline and random seed. What becomes
apparent is that for constant and unsorted distributions the number of completed
workflows is almost linear. This confirms that the budget constraint has the
linear influence on the results, when the influence of the deadline is
eliminated by averaging. On the other hand, for sorted distributions the
functions are convex, increasing slowly for small budgets and steeper for larger
budgets. This reflects the fact that the in sorted ensembles the largest
workflows have the highest priority, so it requires a large increment of a small
budget to fir the next workflow into the schedule.

The chacacteristics discussed in this section can be useful not only for
analysis of our algorithms, but also for planning scientific experiments and
understanding tradeoffs between cost, deadline and the number of work
complete~\cite{Malawski-FGCS-13}.

\begin{figure*}[t!]
    \centering
    \includegraphics[width=0.9\textwidth]{run-finish-variations-test-0-output-deadlines-app-dist-selected}    
    \caption{Number of completed workflows from the ensemble depending on normalized deadline.}
    \label{fig:completed-deadlines}
\end{figure*}

\begin{figure*}[t!]
    \centering
    \includegraphics[width=0.9\textwidth]{run-finish-variations-test-0-output-budgets-app-dist-selected}    
    \caption{Number of completed workflows from the ensemble depending on normalized budget.}
    \label{fig:completed-budgets}
\end{figure*}

\subsection{Task Granularity}
\label{sec:granularity}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]{run-stretching-test-output-scaling-CYBERSHAKE}
    \includegraphics[width=0.9\columnwidth]{run-stretching-test-output-scaling-MONTAGE}

    \caption{Percentage of high scores achieved by each algorithm for CyberShake
    and Montage ensembles when task runtime is scaled by a constant factor. A
    scaling factor of $x$ means that the runtime of tasks in each workflow was
    multiplied by $x$.}
    \label{fig:stretching}
\end{figure}

In the experiments described in previous sections we noted that, for Montage and
CyberShake, the short runtimes of their tasks made the dynamic algorithms
perform better relative to SPSS. In order to test this theory we adjusted the
granularity of the tasks in several Montage and CyberShake ensembles to see how
this would affect the relative performance of the algorithms. The granularity
adjustment was achieved by multiplying the runtime of each task by a fixed
scaling factor.

Fig.~\ref{fig:stretching} shows the relative performance of the algorithms as
the scaling factor is increased from 1 to 16 for CyberShake  and Montage
applications. Each data point in the figure represents 500 simulations (5
ensembles with 10 budgets and 10 deadlines). The best scores percent was
calculated the same way as it was for Fig.~\ref{fig:distributions}.

The figure shows that, as the scaling factor increases beyond 2, the relative
performance of SPSS surpasses that of the dynamic algorithms.  This result
suggests that, in general, for fine-grained workflows the dynamic algorithms
will produce better scores, and for coarse-grained workflows SPSS will produce
better scores.

The granularity of the workflow tasks should be discussed in comparison with the
length of the cloud billing cycle. In this paper we assumed that the resources
are charged for hourly time slots, as is the case of Amazon EC2. However, there
are cloud providers that have higher billing frequency, e.g. every 5 minutes in
the case of CloudSigma\footnote{\url{http://www.cloudsigma.com}} or every 1
minute in the case of Google Compute
Engine\footnote{\url{https://cloud.google.com/pricing/compute-engine}} (after
first 10 minutes). Our results indicate that when the task granularity is close
to the length of the billing period, the SPSS produces better results, while for
tasks that are shorter than the billing period, the dynamic algorithms perform
better. This suggests that the these effects have to be taken into account when
selecting a cloud provider, and also when using task clustering algorithms that
group smaller tasks into larger groups~\cite{ChenD12}.
